\subsubsection{nETL Module}
Packaged as a node.js module, the intention is that a user can simply import the \mintinline{javascript}{NETL} module and instantiate the \mintinline{javascript}{NETL} class as a singleton. Thereafter a user can add their own extraction / transformation / load modules via the singleton's API (the returned object). The format of these modules should conform to the interface as represented in the diagram. (Actually, JavaScript is not a suitable language for specifying interfaces since tooling that allows interface implementation in the language easy is lacking).

As specified in Figure \ref{nETL}, modules should return an object with the properties 'name' and 'exe'. 'name' should be the naming identifier of the function, and 'exe' should return the starting function of the module. Adding modules to the framework makes them available to tasks as specified by configuration objects. These objects are implemented as a JSON, and are used as instructions to start Tasks. Effectively this allows users to specify custom data processing tasks as a configuration + module. The configuration is passed to the program on startup, and the tasks are run. A possible netl startup script with in line configuration and modules may look like this:

\begin{minted}{javascript}
/* File <user entry point>.js */

// Import the nETL module
const NETL = require('./path/to/netl.js');
var netl = NETL();

// Specify the extraction/transformation/load configuration
const config = [{
    "ID": "TaskName",
    "Extraction": {
        "Name": "ExtractionType",
        // ...
        "afterTaskRunCBs": ["function(moduleConfig) {console.log('Run on task-end')}"]
    },
    "Transformations": [{
            "Name": "TransType",
            // ...
            "afterTaskRunCBs": ["function(moduleConfig) {console.log('Run on task-end')}"]
        },
        {
            // ...
        },
        // ...
    ],
    "Load": {
        "Name": "LoadType",
        // ...
        "afterTaskRunCBs": ["function(moduleConfig) {console.log('Run on task-end')}"]
    }
}];

// Load your own bespoke extraction module
netl.loadExtractionModule((function() {
    function exe(obj) {
        // ...
    };
    return {
        name: "ExtractionType",
        exe: exe
    };
})());

// Load your own bespoke Transformation module
netl.loadTransformationModule((function() {
    function exe(obj) {
        // ...
    };
    return {
        name: "TransType",
        exe: exe
    };
})());

// Load your own bespoke Load module
netl.loadLoadModule((function() {
    function exe(obj) {
        // ...
    };
    return {
        name: "LoadType",
        exe: exe
    };
})());

// Run your task till completion
netl.taskManager.newTask(config);
\end{minted}